From Reversible Programs to \\ Univalent Universes and Back

### 2021-05-26 Pi+/paper/reversible.tex

\section{The groupoid of finite types}~\label{sec:finite}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

>>>>

\section{Reversible Programming Languages}~\label{sec:reversible}

\todo{Not the right title.}

\note{This section should explain the main technical parts of the paper
  informally, without using any technology. Use an example, such as, a
  reversible language with $\leq 5$ bits, and examples of permutations and
  transpositions, and when they're equal.}

\note{This might be followed by a section which explains the syntax of Pi.}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% fill-column: 120
%%% End:

### 2021-05-26 Pi+/paper/univalent.tex
Add some explanation about Lehmer codes
Progress on #108 and #109

@@

@@

@@

From the topological viewpoint, a type family can also be seen as a fibration. For a type family $P : A \to \UU$ and a
point $x : A$, $P(x)$ gives the fiber over $a$. For a path $p : x \id_{A} y$, $\transport{P}{p}$ gives the path lifting
operation. The total space is given by $\dsum{x:A}{P(x)}$ and the first projection $\pi_1 : \dsum{x:A}{P(x)} \to A$ to
the base space is the fibration. The lifting operation lifts paths in the base space to paths in the total space. If
$p : x \id_{A} y$ is a path in the base space, and $u : P(x)$, we have

@@

@@

@@

@@

\begin{definition}[Set Quotient]
  Given a type $A$ which is an $\hSet$, and a relation $R : A \to A \to \hProp$, the set-quotient $\quot{A}{R}$ is the
  higher inductive type generated by
  \begin{itemize}
    \item an inclusion function $q : A \to \quot{A}{R}$,
    \item for each $x, y : A$ such that $R(x,y)$, a path $q(x) \id_{\quot{A}{R}} q(y)$,
    \item a set truncation, for each $x, y : \quot{A}{R}$ and $r, s : x \id_{\quot{A}{R}} y$, we have $r \id s$,
  \end{itemize}
  with an appropriate induction principle.
\end{definition}

@@

@@

\begin{definition}[$\Fin$]
  The type family $\Fin : \Nat \to \UU$ is the type of finite sets indexed by cardinality. It is defined equivalently in
  two different ways,
  \begin{gather*}
    \begin{aligned}
      \Fin[n] & \defeq \dsum{k:\Nat}{k \leq n}
    \end{aligned}
    \begin{aligned}
      & \Fin[0] \defeq \bot \\
      & \Fin[\suc[n]] \defeq \top \sqcup \Fin[n]
    \end{aligned}
  \end{gather*}
  Note that $\Fin[n]$ is a $\hSet$, and we use the definitions interchangeably.
\end{definition}

@@

We characterise the path space of univalent subuniverses.

@@

\begin{corollary}
  For every $n:\Nat$,
  \[
    \loopspace[\UFin[n],\Fin[n]] \eqv \Aut[\Fin[n]]
  \]
\end{corollary}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

>>>>

\note{It is important that we work in HoTT using it as a metatheory, we care about proof relevance, constructivity, and
  working with higher-dimensional algebraic structures, such as groupoids. Our results are also formalised using general
  tools and principles from HoTT, and computer checked using a proof assistant, using Agda and the HoTT-Agda library.}

@@

\vc{I'm explaining everything starting from the categorical language. We do care more about the groupoid structure of
  types, and how to encode groupoids using types, since that's what we use in~\cref{sec:finite}.}

@@

\todo{Say why it is a groupoid by listing some groupoid laws.}

@@

\todo{Why is the topological viewpoint important, or saying fibrations? Explain the intuition using classifying spaces,
  maybe draw some diagrams.}

From the topological viewpoint, a type family can also be seen as a fibration. For a type family $P : A \to \UU$ and a
point $x : A$, $P(x)$ gives the fiber over $a$. For a path $p : x \id_{A} y$, $\transport{P}{p}$ gives the path lifting
operation. The total space is given by $\dsum{x:A}{P(x)}$ and the first projection ${\pi_1 : \dsum{x:A}{P(x)} \to A}$ to
the base space is the fibration. The lifting operation lifts paths in the base space to paths in the total space. If
$p : x \id_{A} y$ is a path in the base space, and $u : P(x)$, we have

@@

\todo{Explain motivation.}

@@

\todo{Revise.}

@@

\todo{Explain homotopy types, $\hProp$, $\hSet$, etc.}

\vc{These are placeholder definitions, need informal explanations and references to the book.}

@@

\begin{definition}[Set Quotient]
  Given a type $A$ which is an $\hSet$, and a relation $R : A \to A \to \hProp$, the set-quotient $\quot{A}{R}$ is the
  higher inductive type generated by
  \begin{itemize}
    \item an inclusion function $q : A \to \quot{A}{R}$,
    \item for each $x, y : A$ such that $R(x,y)$, a path $q(x) \id_{\quot{A}{R}} q(y)$,
    \item a set truncation, for each $x, y : \quot{A}{R}$ and $r, s : x \id_{\quot{A}{R}} y$, we have $r \id s$,
  \end{itemize}
  with \review{an appropriate induction principle.}
\end{definition}

@@

\todo{Explain motivation. Starting from a univalent universe which classifies all types, we want to define a subuniverse
  which classifies only certain types, for example, types that satisfy some property that we want.}

@@

\begin{definition}[$\Fin$]
  The type family $\Fin : \Nat \to \UU$ is the type of finite sets indexed by cardinality. It is defined equivalently in
  two different ways,
  \begin{gather*}
    \begin{aligned}
      \Fin[n] & \defeq \dsum{k:\Nat}{k < n}
    \end{aligned}
    \begin{aligned}
      & \Fin[0] \defeq \bot \\
      & \Fin[\suc[n]] \defeq \top \sqcup \Fin[n]
    \end{aligned}
  \end{gather*}
  Note that $\Fin[n]$ is a $\hSet$, and we use the definitions interchangeably.
\end{definition}

@@

\review{We characterise the path space of univalent subuniverses.}

@@

\begin{corollary}
  For every $n:\Nat$,
  \[
    \loopspace[\UFin[n],F_{n}] \eqv \Aut[\Fin[n]]
  \]
\end{corollary}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% fill-column: 120
%%% End:
